---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
attach(brfss2013)
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Research quesion 1:**
To plot the number of people who does not have any health coverage plan and could not consult a doctor due to financial isuues, are suffereing from any type of cancer.

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r to find cancer patients}
no_cancer_pateint <- brfss2013 %>%
  group_by(genhlth)%>%
  filter(hlthpln1 == "No", medcost == "Yes")
+ mutate(chcocncr = ifelse(chcocncr == "Yes", "Has cancer", "Dont have cancer"))
ggplot(data = cancer_pateint, aes(x= chcocncr))
+ geom_histogram()
  
```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

